---
title: "Cleaning the Observed and Thinking About the Unobserved Data"
author: "Iiro Tiihonen"
format: revealjs
---

```{r echo = FALSE}
library(magrittr)
library(igraph)
library(ggraph)
```

## Goals of this lecture

-   This lecture will introduce basic problems that affect all practical work with (network) data
-   By the end you should:
    -   Know why missing data matters
    -   Know why unharmonised data matters
    -   Be working with the data set you acquired last time

## Missing data

-   Two major variants:
    -   Partially missing data (known unknown)
    -   Completely missing data (at worst, unknown unknown)

## Partially Missing Data - Three problems:

-   NA is not a number

-   The way the first problem is solved can affect the results

-   Survival bias \## Partially missing data - example

```{r echo=FALSE}

example_data_set <- matrix(nrow=2,ncol=2) %>% as.data.frame(.)
colnames(example_data_set) <- c("name","occupation")
example_data_set[1,1] <- "Thomas Hobbes"
example_data_set[2,1] <- "Steven Steel"
example_data_set[1,2] <- "philosopher"
print(example_data_set)

```

## Partially missing data - common solutions

-   Disregard observations with missing data (e.g. nodes without birth year)\

-   Impute missing values (e.g. fill the missing values with mean, prediction from a statistical model etc.)

-   Robust analysis/analysis of the sensitivity of the results

-   Unfortunately, the first two solutions can be dangerous and affect the results of analysing the data significantly

-   Positive note: Here substance understanding about humanities can be valuable 


## Completely missing data

-   Some nodes (e.g. persons, books) might be missing entirely from the data

-   Reasons vary (failures in querying, survival of historical evidence, etc.)

-   The questions remains: is the data representative or complete enough that analysing it is meaningful

-   Think carefully how important possible missing data is for you research 

## Completely missing data - some solutions

-   Comparison of the data used to other sources

-   Does subject understanding prove insight to the nature of missingness (e.g. book history)

-   Sensitivity analysis of results (how much to they wary if you "shake" your data)

-   A very important topic, but we can not dwell deep into it during the course. 

## Unharmonised data

-   Definition of unharmonised: things that are exactly same to us and should be considered the same, are not the same in the data

-   Example: Names of publishers in the ESTC with spelling variations

-   Solution: harmonisation

## Unharmonised data - example

```{r echo=FALSE}
aachen <- c("Aach","Achen", "Aix la Chapelle", "Aix-la-Chapelle")
print("Some name variants of the City of Aachen according to CERL Thesaurus:")
print(aachen)
```

## Unharmonised data - why it matters

-   Very common phenomenon in all sorts of data sets, especially with text data
-   Without harmonisation, many datasets - for example networks - will be heavily inflated

## Unharmonised data - common solutions

-   Start by just taking a look at your data, can you e.g. notice the same name being spelled in different ways?
-   Most common solutions:
    -   Via programming: regular expressions (systematic changes to character strings)
    -   Most of the book chapter is about using regular expressions with Stringr
    -   With software designated to the task (e.g. openRefine)
    
## Unharmonised data - we focus on one tool out of many    
-   A field in its own right, all sorts of algorhitms and metrics have been developed to harmonise data.
-   We will practice with regular expressions, they are flexible, powerful and relatively quick to learn

## Summary

-   Always consider these three questions when working with (network) data
    -   Are you completely missing some of the relevant data, and can you evaluate the severity of this problem?
    -   What will you do to your partially missing data (imputation, removal of nodes with missing data, do the results of your analysis depe´ your data and try to find if there is a need to harmonise it (always assume that this is the case).

## What Next
- Start working on the exercises and going through the related book chapter
- Remember also the article for the next week, though its better to use
remainder of the lecture for the programming task.