---
title: "Introduction to R and the Tidyverse"
output:
  html_document:
    df_print: paged
---

## EXERCISES - R Basics and Tidyverse

Change the third item in the `fruit` vector to `bananas` and the corresponding item in the colour vector to `yellow`:

```{r}
fruit = c("apples", "bananas", "oranges", "apples")
colour = c("green", "yellow", "orange", "red")
amount = c(2,5,10,8)
```

Turn this into a dataframe:

```{r}

```

Get a total for the amount of each fruit:

```{r}

```

Read in the `book_sample.csv` dataset, giving it the name `df`. It's a random sample of 1500 books published in London, Edinburgh, and Dublin. You'll need to find this file in the 'data' folder of the CSC notebooks workspace.

Fill in the necessary part of the function below:

```{r}

```

Filter to only include books with Dublin as the place of publication. Call this new dataset `dublin_books`:

```{r}


```

Count the number of books per year in the `dublin_books` dataset:

```{r}

```

Count the number of books for each place of publication in the original `df` dataset. Save this as an object called `df_count`.

```{r}

```

Arrange the places in descending order of number of books:

```{r}

```

Use `ggplot2` to make a bar chart of the count of places:

```{r}

```

Join the `df_count` dataset to the `df` dataset, so that each record now has the total count alongside it.
