
---
title: "Applying Network Analysis to Humanities: Week 2"
subtitle: "Data Harmonisation"
output:
  html_document:
    df_print: paged
---

The exercise of this lesson uses regular expressions to solve a real data wrangling problem with the Tweetsets data. Run the chunk below so that you 
can work with the Tidyverse functions for the rest of the exercise 
```{r}
library(tidyverse)
```
## Warm up
It is a good idea to first practice with simple examples. Do that in the code chunk below, by filling str_example with content and then extracting, replacing or otherwise modifying it with some str_function, and saving the output of that modification to str_example_modified. It is a good idea to test many things, but after you feel you have a grasp of the basics of regular expressions and str functions, save an example string to str_example, and an example modification to str_example_modified.
```{r}
str_example <- "Your practice string here"
str_example_modified <- "use one of the str functions to replace, extract or otherwise manipulate the str_example, and save the output to this variable" 
print(str_example_modified)
```

## Regular expression excercise
When writing the material for this course, we noticed that the Twitter data that Tweetsets returns gives the dates in the following format (run the chunk, we will 
need this string):
```{r}
date_example <- "Fri Apr 06 01:28:26 +0000 2018"
```

We wanted to have this information in the format day, month, year, so we used regular expressions to extract
the information, and other R functions to turn it to the date data format. In fact, we made
a little function of our own to do this. Now, you will replicate this step.

If defining a function in R is a new thing to you, don't worry, we keep it simple. Function takes something as input (in our case the string x), and returns something as an output (defined by the line return, that should be the last line in a simple function like this). Self-made function is wirst defined, after that you can use it like any of the R functions you are already using like count or str_extract. 

Here is a small example to get the idea of how self defined functions work when combined with the other tools that we have discussed. 

```{r}
#Function starts  by first naming it, defining the input (in this example by writing function(x),and writing a bracket {. 
#Function definition begins here
friday_detector <- function(x){
  #This function takes the argument x, converts it to lower case, then detects "fri" from it, and replaces it
  #with the following text: friday was removed from here.
  x_without_friday <- x %>% tolower(.) %>% str_replace(.,pattern = "fri",replacement = "friday was removed from here")  
  #return x_without_friday as the output of the function
  return(x_without_friday)
#Function ends to a bracket }
}
#Function definition ends here

#The output of the function is saved to use_function_demo
#IMPORTANT: now that date_example is the argument x given to the function, the function performs is task (detecting and removing friday) from date_example.
use_function_demo <- friday_detector(date_example)
print(use_function_demo)

```
As you can see, after running the code chunk above, we are able to use the function like any of the pre-made functions of R that you have already used. You can modify the function in the chunk above to get a touch on how functions (and regular expressions) work. 

Next, you should finalise the function polish_to_date. As an input (x), the function is given a character vector of time information in the same format as in our date_example vector (so time is expressed in this order and way of writing it: "Fri Apr 06 01:28:26 +0000 2018"). It is your task to use regular expressions to extract the day and month (e.g jun 29, may 20) to the day_and_month variable of the function and year (e.g. 2018) to the year variable. We have already started by
converting the input (a string with time information in the same structure as the string date_example) to lower case. Your should continue by piping the output forward to a str function (which one of them, you have to figure out) in both lines. Try to find the right function and pieces of regular expression either from the cheat sheet or from the course book.

```{r}
polish_to_date <- function(x){
# Pipe the x put to lower case forward, and use right str-function and regular expression to get the day and month to the variable day and month
day_and_month <- x %>% tolower(.) 
# Pipe the x put to lower case forward, and use right str-function and regular expression to get the year to the variable year
year <- x %>% tolower(.)
#  You don't need to touch the date variable
date <- paste(day_and_month,year,sep=" ") %>% as.Date(.,format="%b %d %Y")
return(date)

}

```
The chunk below should return the output 2018-04-06 if the function works as it should. You should be able to use the function as
such to harmonise the time information (created_at variable) of your Tweetsets data.
```{r}
harmonised_date <- polish_to_date(date_example)
print("Date harmonised:")
print(harmonised_date)

```
Finally, return the rmd and html files to Moodle.
