---
title: "Visualising Networks with ggraph"
output:
  html_document:
    df_print: paged
---

## Exercises - Visualising your data with ggraph.

### Make the network object.

Using the `letter_data1670.csv` dataset, create a network object. You can do each step in turn in the cells below. 
Load the libraries `igraph`, `tidygraph`, `tidyverse` and `ggraph`:

```{r}
library()
```

Load in the file using `read_csv()` and save it as an object called `letter_data`:

```{r}

letter_data = 

```


Count the number of letters between each pair of `from_id, to_id`, as a weight, and it as an object called `edge_list`


```{r}
edge_list = letter_data %>% 
  group_by()
```

Turn the edge list into a network object, called `g`, with `as_tbl_graph()`

```{r}
g = edge_list %>% 
```


In the following cells, when we make a set of commands to create a visualisation, we won't save it as a new object. Instead, we need to repeat the same set of commands each time. 

So for example to draw a basic points and lines graph we use the following code: 

```{r}
g %>% ggraph() + geom_edge_link() + geom_node_point()
```

And then to change the layout method to the 'circle' we repeat the full code but make a change where appropriate:  

```{r}
g %>% ggraph(layout = 'circle') + geom_edge_link() + geom_node_point()
```

For the following commands you should do the same - copy and paste your existing code into the new cell, and make the changes. 

Draw the graph with the 'Fruchterman-Reingold' layout:

```{r}

```

Change the size of the nodes to the value 5:

```{r}
g %>% 
```

-   Calculate the betweenness centrality of the nodes in the network, and save it as a column using `mutate(). 

-   Filter to only include nodes with a betweenness centrality score of over 5. 

-   Map the betweenness score to the size aesthetic, for the nodes in your network diagram:

```{r}

```

Filter to only include edges with a weight of greater than 1, and again draw the network. You'll need to use the `activate()` command at the correct point to do commands on the edges table in the `tbl_graph` object. 

```{r}

```

## Extra exercises (for 'fun')

Please only do if you are confident, or go through these earlier ones quickly!

1. 

One advantage of doing network analysis in R is that you can do complicated data wrangling steps and send the results straight to a network. 

Join the `node_attributes.csv` file we used in last week's lesson to the `tbl_graph` object. Filter the data to only include rows where the `main_name` contains the word 'Sir', and draw this as a network:

```{r}

```

2. Last one (warning: you'll have to use external resources to figure it out!)

We can also use functions to show certain aesthetics using conditional statements - such as 'if' something is true, then draw the aesthetic, but otherwise leave it empty. 

With the network, join the node attributes again, and calculate the degree score. Add a geom which draws a label for each node. This label should contain the `main_name` for each node. 

Within this geom, add a conditional statement which says: if the degree score for this node is greater than 10, create the label `main_name`, otherwise, leave it as `NA`. 

The R command `ifelse()` is a starting point.. but you'll have to figure out where in the process to put it. 

```{r}

```


