---
title: "R Notebook"
output: html_notebook
---

In your group, design a network visualisation of one of the sets of network data we have used over the past few weeks.

Designate one person as the coder. This person should share their screen with the group and enter the code for the network in the notebook below. 

You can use the course book to find code examples to visualise different elements of your network data. 

Think about how you can use ggraph to make the network more readable and communicate your desired information to the person viewing it. 

-   Experiment with the layouts, taking into account your particular network
-   Map values to aesthetics, such as size, shape and transparency
-   Filter or pre-process your data - to make a large network more readable, for example. 

Be prepared to present your network to the rest of the groups, and talk through briefly why you made your design decisions. 

Run this cell to load libraries you'll need:

```{r}
library(tidyverse)
library(tidygraph)
library(igraph)
library(ggraph)
```


Potential datasets - run the cell to load and use that particular one. 

17th century letter data and related node attributes:  

```{r}
letter_data = read_csv('../data/letter_data1670.csv')

node_attributes = read_csv('../data/node_attributes.csv')

```

Book publisher network, taken from the English Short Title Catalogue (ESTC):

```{r}
publisher_network = read_csv('../data/sample_publisher_network.csv')
```

Small sample of a twitter edge list, using Tweetsets: here, the tweets are from the 'Repeal the 8th' campaign in Ireland (https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/PYCLPE).

```{r}
repeal_edge_list = read_csv('../data/repeal_edge_list_sample.csv')
```


